---
title: "distributions_of_final_data"
output: html_document
date: "2023-08-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data and libs}
library(knitr)
library(dplyr)
library(tidyverse)
library(gridExtra)
data <- read.csv('./our_data/new_stops.csv')
```

# Looking at the distributions of the final drivers dataset 
### Average distance per stop
```{r average distance per stop}
data %>% ggplot(aes(x=average_distance_per_stop)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density of Average Distance per Stop")
```

Super NB: look at that weird spike in the negatives. Should investigate.
Let me look at it as a boxplot:

```{r average distance per stop box}
data %>% ggplot(aes(x=average_distance_per_stop)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Density of Average Distance per Stop Boxplot")
```
Let's filter out that bottom neg values:
```{r average distance per stop box filter}
plot1 <- data %>% filter(average_distance_per_stop > 0) %>% ggplot(aes(x=average_distance_per_stop)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot of Average Distance", x = "avg distance per stop")
plot2 <- data %>% filter(average_distance_per_stop > 0) %>% ggplot(aes(x=average_distance_per_stop)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density of Average Distance", x = "avg distance per stop")
grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```
Now that makes a lot more sense wow.

### Average number of Stops per day
```{r average nmr of stops per day}
data %>% ggplot(aes(x=average_number_of_stops_per_day)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density of Average Number of Stops per day", x = "avg number of stops per day")
```
Also NB: shoud look at that weird spike near 800 by making a boxplot of this dude

```{r average nmr of stops per day box}
data %>% ggplot(aes(x=average_number_of_stops_per_day)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot of Average Number of Stops per day", x = "avg number of stops per day")
```
Ooh la, la. Look, that is a SINGLE outlier near 800. Let's remove that dude as well. However, we should 100% investigate !!

## Looking at the average nmr of stops per day without outliers
```{r average nmr of stops per day filtered box}
plot1 <- data %>% filter(average_number_of_stops_per_day < 700) %>% ggplot(aes(x=average_number_of_stops_per_day)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot", x = "avg number of stops per day")
plot2 <-  data %>% filter(average_number_of_stops_per_day < 700) %>% ggplot(aes(x=average_number_of_stops_per_day)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density", x = "avg number of stops per day")

grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))

```
Now, those look tons better.

### Looking at the correlation between some things
```{r freq over speed vs total claims cost}
data %>% filter(average_number_of_stops_per_day <700)  %>% ggplot(aes(x = average_number_of_stops_per_day, y = total_claims_cost)) + geom_jitter(color= "blue", alpha=0.8) + labs(title="Average nmr of stops per day vs Total Claims Cost", x= "Average nmr of stops per day", y= "Total Claims Cost") + geom_smooth(color = "grey")
data %>% filter(average_number_of_stops_per_day <700)  %>% ggplot(aes(x = average_number_of_stops_per_day, y = number_of_claims)) + geom_jitter(color= "blue", alpha=0.8) + labs(title="Average nmr of stops per day vs Number of Claims", x= "Average nmr of stops per day", y= "Number of Claims") + geom_smooth(color = "grey")
```

### Total claims cost

```{r total claims cost}
plot1 <- data %>% ggplot(aes(x=total_claims_cost)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density of Total claims Cost", x = "Total claims Cost (R)")
plot2 <- data %>% ggplot(aes(x=total_claims_cost)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot of Total claims Cost", x = "Total claims Cost (R)")

grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```
Let us inspect that Boxplot a little more. I feel like the real density lies beyond all those 0's.


```{r total claims cost filtered}
data %>% filter(total_claims_cost > 0) %>% ggplot(aes(x=total_claims_cost)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot of Total claims Cost", x = "Total claims Cost (R)")
```
A-Hah! Now that seems to look like a density. Look at that one dude wth R4 0000 +, that is sort of hectic. 

### Over speed limit

```{r overspeed}
plot1 <- data %>% ggplot(aes(x=overSpeed)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density of Over Speed", x = "Over Speed")
plot2 <- data %>% ggplot(aes(x=overSpeed)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot of Over Speed", x = "Over Speed")

grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```
We should probably go look at those guys that have 9000x over the speed limit

```{r freq over speed vs total claims cost}
data %>% ggplot(aes(x = overSpeed, y = total_claims_cost)) + geom_jitter(color= "blue", alpha=0.8) + labs(title="Over Speed freq vs Total Claims Cost", x= "Times over speed limit", y= "Total Claims Cost") + geom_smooth(color = "grey")
```
Ok, but those total_claims = 0 are annoying things. Let's filter them out. 

```{r freq over speed vs total claims cost filter}
data %>% filter(total_claims_cost > 0) %>% ggplot(aes(x = overSpeed, y = total_claims_cost)) + geom_jitter(color= "blue", alpha=0.8) + labs(title="Over Speed freq vs Total Claims Cost", x= "Times over speed limit", y= "Total Claims Cost") + geom_smooth(color = "grey")
```
Ok, this is still weird. I don't think this actually has a proper effect. 

## Normal Braking 

```{r normal braking}
plot1 <- data %>% ggplot(aes(x=normal_braking)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density Normal Braking", x = "Normal Braking")
plot2 <- data %>% ggplot(aes(x=normal_braking)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot Normal Braking", x = "Normal Braking")
grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```

## Emergency Braking 
```{r emergency braking}
plot1 <- data %>% ggplot(aes(x=emergency_braking)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density Emergency Braking", x = "Emergency Braking")
plot2 <- data %>% ggplot(aes(x=emergency_braking)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot Emergency Braking", x = "Emergency Braking")
grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```
## Harsh Braking 
```{r emergency braking}
plot1 <- data %>% ggplot(aes(x=harsh_braking)) + geom_density(fill="lightblue", color="lightblue") + labs(title="Density Harsh Braking", x = "Harsh Braking")
plot2 <- data %>% ggplot(aes(x=harsh_braking)) + geom_boxplot(fill="lightblue", color="black") + labs(title="Boxplot Harsh Braking", x = "Harsh Braking")
grid.arrange(plot1, plot2, ncol = 2, widths = c(3, 3))
```
```{r summarize the braking data}
data %>% select(harsh_braking, normal_braking, emergency_braking) 
```
